### 2018-04-25 General WLD combinatorics.tex
finished completing Sian's proof that the GN is correct

To prove that this algorithm gives the Grassmann necklace of the diagram we need a lemma.

@@

\begin{prop}\label{res:alg gives GN}
The sequence of $k$-subsets obtained by applying Algorithm~\ref{alg:put GN on WLD} to an admissible diagram $W$ is exactly the Grassmann necklace of the positroid associated to $W$.
\end{prop}

\note{I think the statement should be clearer about what this sequence is, in the way that the first line of the proof currently is.}

@@

\note{I tried to make the first part of this proof a bit tidier by using contradiction.  I will do the same for the second part (what is currently below) and also finish the proof, when I have a chance}

We have shown that $\II$ is a Grassmann necklace; it remains to check that this Grassmann necklace defines the positroid $\PP_W$ associated to $W$.  We need to show that:
\begin{itemize}
\item For each $i \in \interval{1}{n}$, $I_i$ is a basis for $\PP_W$.
\item If $J$ is lexicographically smaller than $I_i$ with respect to $<_i$, then $J$ is not a basis for $\PP_W$.
\end{itemize}
Recall from [ref] that a $k$-subset $J \in \binom{[n]}{k}$ is a basis for $\PP_W$ if and only if it has no subset $U \subseteq J$ such that $|U| > |Prop(U)|$.  It is immediately clear from the construction that $I_i$ is a basis for each $i$: the algorithm is effectively pairing each $j \in I_i$ with a unique propagator supported on that vertex, so $|Prop(U)| \geq U$ for all $U \subseteq I_i$.

If $\exists l \in J$ with $Prop(l) = \emptyset$ then $J$ is clearly not a basis, so suppose otherwise.  Write
\[I_i = [i_1 <_i i_2 <_i \dots <_i i_r <_i \dots <_i i_k], \quad J = [i_1 <_i i_2 <_i \dots <_i j_r <_i \dots <_i j_k],\]
so that $I_i$ and $J$ differ for the first time in the $r$th position.  Then
\[r-1 \geq |Prop(\{i_1, \dots, i_{r-1}\})| = |Prop(\{i_1, \dots, i_{r+1},j_r\})|,\]
If $|Prop(\{i_1, \dots, i_{r-1}|\}| = r-1$ then we are done; otherwise

>>>>

To prove that this algorithm gives the Grassmann necklace of the diagram we need two lemmas.

@@

\begin{lem}\label{lem basis as perm}
Let $W$ be an admissible Wilson loop diagram and let $\PP_W$ be the positroid associated to $W$.  A $k$-subset $J \in \binom{[n]}{k}$ is a basis for $\PP_W$ if and only if there is a set bijection between $J$ and the set of propagators of $W$ such that for each $j\in J$ the propagator associated to $j$ is supported on vertex $j$.
\end{lem}

\begin{proof}
\note{I don't know if you already have this in your previous papers or not; but it is important so I wanted to extract it as a lemma.  The proof of the next proposition originally contained the lines ``Recall from [ref] that a $k$-subset $J \in \binom{[n]}{k}$ is a basis for $\PP_W$ if and only if it has no subset $U \subseteq J$ such that $|U| > |Prop(U)|$.'' but I don't know what that ref is nor whether that ref would also prove this lemma.}
\end{proof}

\begin{prop}\label{res:alg gives GN}
The sequence of $k$-subsets obtained by applying Algorithm~\ref{alg:put GN on WLD} to an admissible diagram $W$ is exactly the Grassmann necklace of the positroid associated to $W$; specifically if we let  $I_i$ be the set of labels on edge $(i,i+1)$ then the sequence $\II = (I_1, \dots, I_n)$ is the Grassmann necklace of the positroid associated to $W$.
\end{prop}

@@

\medskip

We have shown that $\II$ is a Grassmann necklace; it remains to check that this Grassmann necklace defines the positroid $\PP_W$ associated to $W$.  We need to show that:
\begin{itemize}
\item For each $i \in \interval{1}{n}$, $I_i$ is a basis for $\PP_W$.
\item If $J$ is lexicographically smaller than $I_i$ with respect to $<_i$, then $J$ is not a basis for $\PP_W$.
\end{itemize}
The algorithm is pairing each $j \in I_i$ with a unique propagator supported on that vertex so by Lemma~\ref{lem basis as perm} $I_i$ is a basis for $\PP_W$.
%%Recall from [ref] that a $k$-subset $J \in \binom{[n]}{k}$ is a basis for $\PP_W$ if and only if it has no subset $U \subseteq J$ such that $|U| > |Prop(U)|$.  It is immediately clear from the construction that $I_i$ is a basis for each $i$: the algorithm is pairing each $j \in I_i$ with a unique propagator supported on that vertex, so $|Prop(U)| \geq U$ for all $U \subseteq I_i$.

%%If $\exists l \in J$ with $Prop(l) = \emptyset$ then $J$ is clearly not a basis, so suppose otherwise.  Write
%%\[I_i = [i_1 <_i i_2 <_i \dots <_i i_r <_i \dots <_i i_k], \quad J = [i_1 <_i i_2 <_i \dots <_i j_r <_i \dots <_i j_k],\]
%%so that $I_i$ and $J$ differ for the first time in the $r$th position.  Then
%%\[r-1 \geq |Prop(\{i_1, \dots, i_{r-1}\})| = |Prop(\{i_1, \dots, i_{r+1},j_r\})|,\]
%%If $|Prop(\{i_1, \dots, i_{r-1}|\}| = r-1$ then we are done; otherwise

Suppose we have $J$ such that $J$ is a basis and yet is lexicographically less than $I_i$ with respect to $<_i$.  By Lemma~\ref{lem basis as perm} there is a set bijection between $J$ and the propagators of $W$ such that the propagator associated to $j$ is supported on vertex $j$.  Choose one such bijection.  For a propagator $p$ of $W$ write $J(p)$ for the associated $j$ according to this bijection.  Similarly write $I_i(p)$ for the vertex assigned to $p$ by the algorithm.  Since $J$ is lexicographically smaller than $I_i$, the $<_i$-smallest element of the symmetric difference of $J$ and $I_i$ is some $j_0\in J$, $j_0 \not\in I_i$.

Let $p$ be the propagator such that $J(p)=j_0$.  Since $j_0\not\in I_i$ but $p$ is supported at $j_0$, then $I_i(p) <_i j_0$.  Thus the propagator $p$ has the following property: $I_i(p) <_i J(p)$ and $I_i\cap \interval{i}{I_i(p)} = J\cap \interval{i}{I_i(p)}$.  Call this property $A$.

{}From the previous paragraph we conclude that if we ever had a $J$ which is lexicographically less than $I_i$ and yet a basis of $\PP_W$ then there exists a propagator $p$ which has property $A$.

Finally, we will prove that whenever there is a propagator $p$ which has property $A$ then there is another propagator $r$ which also has property $A$ and for which $I_i(r) <_i I_i(p)$.  Since $I_i$ is finite this would lead to infinte regress and hence is impossible and thus there can be no such $J$ which will complete the proof of the proposition.

So suppose $p$ has property $A$.  The same vertices in $\interval{i}{I_i(p)}$ are in the image of the assignment from $J$ and the image of the assignment from $I_i$ so the same number of propagators are assigned to vertices in this interval by each assignment.  However $p$ is one of the propagators assigned to a vertex in this interval in $I_i$ but not in $J$.  Thus there exists a propagator $q$ for which $J(q)\in \interval{i}{I_i(p)}$ but $I_i(q) >_i I_i(p)$.  Therefore $J(q)<_i I_i(q)$ and  $I_i\cap \interval{i}{J(q)} = J\cap \interval{i}{J(q)}$.  This is analogous to property $A$ but with the roles of $J$ and $I_i$ switched.  Thus by the same argument we must have a propagator $r$ for which $I_i(r)\in \interval{i}{J(q)}$ but $J(r)>_i J(q)$.  Therefore $r$ has property $A$ which is what we wanted to prove.
